
### 2020 Raw Data Cleaning ###
### 1/12/2023 ###

### Step 1: Install Packages ###

install.packages("tidyverse")
install.packages("readxl")
install.packages("janitor")

library(tidyverse)
library(readxl)
library(janitor)

### Step 2: Import Data ###
## The readxl package allows you to pull data straight from an excel file instead of having to make a CSV file first ##

```{r}
Spring_2020_raw <- read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx")
## multiple sheets are in this file so we need to specify which sheet we want to work with

Spring_2020_raw <- read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1")
## there are a lot of error messages due to the sheet not being cleaned up properly ##

Spring_2020_raw <- read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", skip = 1)
## the skip line allows you to get rid of a column of the dataset. We want to keep the date so we need to get rid of the skip line ##

# Instead lets perhaps skip naming columns and just have everything come in as characters. 

Spring_2020_raw <- read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = FALSE)
```

### There is more than one year's worth of data however. We need to import all three and check if they're in the same format.###

```{r}
Spring_2020_headers <- 
  read_excel("Spring 2021 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE)

Spring_2021_headers <- 
  read_excel("Spring 2021 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE)

Spring_2022_headers <- 
  read_excel("Spring 2022 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE)

names(Spring_2020_headers)
names(Spring_2021_headers)
names(Spring_2022_headers)
```

Formatting one data frame - let's check out the checklist!

Which row has the best column names to keep?
Use colnames(data) <- data[rownumber, ] to name your columns after that row.

```{r}

# the function row_to_names() is from the janitor package
Spring_2020_raw <- Spring_2020_raw %>% row_to_names(row_number(1))

Spring_2020_raw
```

Are those column names okay, or do they need to be changed to be tidy, contain identifying information? 
Use rename() to give your columns better names. 
## rename(Name you want it to be = 'current column name') ##
```{r}
Spring_2020_raw <- Spring_2020_raw %>% rename(Transect = `Transect #`, Hits = `# Hits`)
```

### Cleaning Up Data ###
## I am starting with cleaning the block and plot columns ##
## I need to change the values from 1.0 to 1 ##

# to make it a factor: 
```{r}
glimpse(Spring_2020_raw) # Block is a character, we want it as a factor
Spring_2020_raw %>% mutate(Block = as.factor(Block))
```

