---
title: "DroughtNetReformatting"
author: "Phyl"
date: "2023-01-12"
output: html_document
---

```{r}
library(tidyverse)
library(janitor)
install.packages("janitor")
library(readxl)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
#importing data
#readxl package will help with importing data form XLS or xLSX

#Spring_2020_raw <- read_excel ("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet =2)

#View(Spring_2020_raw)

```
#need to know how to import various documents of data that are potentially in different formats
```{r}
Spring_2020_raw <-
  read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = FALSE)

Spring_2021_raw <-
  read_excel("Spring 2021 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = FALSE)

Spring_2022_raw <-
  read_excel("Spring 2022 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = FALSE)

Spring_data_raw <- bind_rows(Spring_2020_raw, Spring_2021_raw, Spring_2022_raw)

Spring_data_raw_better <- Spring_data_raw %>%
  row_to_names(row_number(1)) %>%
  rename(Transect = `Transect #`, Hits = `# Hits`) %>% 
  filter(Date != "Date" & Date != "mm/dd/yy")

view(Spring_data_raw_better)

#get rid of non-data rows
```

#tidying data
```{r}
Spring_data_raw_better %>%
  mutate(Date=as.Date(as.numeric(as.character(Date)), origin = "1899-12-30"),
         Block =as.factor(Block),
         Plot = as.factor(Plot),
         Treatment = as.factor(Treatment),
         Transect = as.factor(Transect),
         Region = as.factor(Region),
         Hits = as.numeric(Hits),
         Distance = as.numeric(Distance)) >%>
  full_join(DNCodes, by = "Code") >%>
  left_join()
         
         
Spring_data_raw_better$Date >%> unique()
#OR
Spring_data_raw_better$Date >%> write.csv(file="DNCodes_raw")

Spring_data_raw_better >%> filter(is.na(Date) == TRUE)

Spring_data_raw_better >%> filter(is.na(Date)== FALSE)
#date, block, plot, treatment, distance, region, transect,

```

#will now need to clean the data set incase there work typos input
```{r}
#use the join function to make sure species are categorized correctly
#start by renaming the following.

Full_Species_List_Glossary_reference_only>%> View()

SpeciesKey <- Full_Species_List_Glossary_reference_only_ >%>
  select(1:7) >%>
  rename(Species Code = 'Species Code (A-Z)',
         NativeStatus = ' Native/Non-Native',
         CommonName = 'Common Name',
         SciName = 'Full Scientific name (Genus species)')

names(Full_Species_List_Glossary_reference_only_)
#^ to help save time with typing
```

join function
```{r}
intermediate <- Spring_data_raw_better >%>
  filter(is.na(Date) == FALSE) >%>
  full_join(DNCodes, by = "Code" ) %>%
  left_join(SpeciesKey, by = " SpeciesCode") >%>
  Lifeform = case_when(n # this column came in from our specieskey and is more complete)
                       is.na(LifeForm) == FALSE ~ LifeForm, #when the vsion from dncodes ha a value substitue int that value to the more complete map
                       TRUE ~as character(Lieforms)),CommonName = case_when(
                         SpeciesCOde == "L" == "itter"
                         Species Code == "B" ~    "bare ground"
                     Species Code =="W" == Coarse woody debris"
                        Species Code == "F"
                         Species Code == "R"
                         Species Code == T == "thatch
                        TRUE ~ as.character(CommonName)
                        
                      SciName = case_when (
                        Type = "Gound" ~ Common names,
                        TRUE ~ as.character(SciName))
```

```{r}

```


Finish this later: see picture on phone
```{r}
tidying:
  Spring_data_raw_better<- Spring_data_raw >%>
  row_to_names(row_number1)) >%> 
  rename(Transect = 'Transect #', Hit = '# Hits') >%>
  )

```

```{r}

```

```{r}
# the function row_to_names()is from the janitor package
Spring_2020_raw <- Spring_2020_raw %>% row_to_names(row_number(1))

Spring_2020_raw

#same as

row_to_names(Spring_2020_raw, row_number = 1)

```

are those column names okay, or do they need to be changed to be tidy, contain identifying information? use rename() to give your columns better names.
```{r}
Spring_2020_raw <- Spring_2020_raw %>% rename(`Transect#`, `#Hits`)
```

Which rows dont contain observations?
remove them from your dataset using [,] or filter()
#as.Date(42705(colname), origin = ")
```{r}
Spring_2020_raw %>% filter("Date" & Date != "mm/dd/yy")
```
# equals is ==
#Greater than is > or lesser than is <
#equal to or greater is =>, and vice versa =<
#doesnt equal is !=
#and is &
#or is |

What is a factor? regarding anova's, it would be a category.
```{r}

```

