<<<<<<< HEAD
---
title: "DroughtNetReformatting"
author: "Phyl"
date: "2023-01-12"
output: html_document
---

```{r}
library(tidyverse)
library(janitor)
install.packages("janitor")
library(readxl)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
#importing data
#readxl package will help with importing data form XLS or xLSX

#Spring_2020_raw <- read_excel ("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet =2)

#View(Spring_2020_raw)

```
#need to know how to import various documents of data that are potentially in different formats
```{r}
Spring_2020_raw <-
  read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = FALSE)

Spring_2021_raw <-
  read_excel("Spring 2021 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = FALSE)

Spring_2022_raw <-
  read_excel("Spring 2022 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = FALSE)

Spring_data_raw <- bind_rows(Spring_2020_raw, Spring_2021_raw, Spring_2022_raw)

Spring_data_raw_better <- Spring_data_raw %>%
  row_to_names(row_number(1)) %>%
  rename(Transect = `Transect #`, Hits = `# Hits`) %>% 
  filter(Date != "Date" & Date != "mm/dd/yy")

view(Spring_data_raw_better)

#get rid of non-data rows
```

#tidying data
```{r}
Spring_data_raw_better %>%
  mutate(Date=as.Date(as.numeric(as.character(Date)), origin = "1899-12-30"),
         Block =as.factor(Block),
         Plot = as.factor(Plot),
         Treatment = as.factor(Treatment),
         Transect = as.factor(Transect),
         Region = as.factor(Region),
         Hits = as.numeric(Hits),
         Distance = as.numeric(Distance)) >%>
  full_join(DNCodes, by = "Code") >%>
  left_join()
         
         
Spring_data_raw_better$Date >%> unique()
#OR
Spring_data_raw_better$Date >%> write.csv(file="DNCodes_raw")

Spring_data_raw_better >%> filter(is.na(Date) == TRUE)

Spring_data_raw_better >%> filter(is.na(Date)== FALSE)
#date, block, plot, treatment, distance, region, transect,

```

#will now need to clean the data set incase there work typos input
```{r}
#use the join function to make sure species are categorized correctly
#start by renaming the following.

Full_Species_List_Glossary_reference_only>%> View()

SpeciesKey <- Full_Species_List_Glossary_reference_only_ >%>
  select(1:7) >%>
  rename(Species Code = 'Species Code (A-Z)',
         NativeStatus = ' Native/Non-Native',
         CommonName = 'Common Name',
         SciName = 'Full Scientific name (Genus species)')

names(Full_Species_List_Glossary_reference_only_)
#^ to help save time with typing
```

join function
```{r}
intermediate <- Spring_data_raw_better >%>
  filter(is.na(Date) == FALSE) >%>
  full_join(DNCodes, by = "Code" ) %>%
  left_join(SpeciesKey, by = " SpeciesCode") >%>
  Lifeform = case_when(n # this column came in from our specieskey and is more complete)
                       is.na(LifeForm) == FALSE ~ LifeForm, #when the vsion from dncodes ha a value substitue int that value to the more complete map
                       TRUE ~as character(Lieforms)),CommonName = case_when(
                         SpeciesCOde == "L" == "itter"
                         Species Code == "B" ~    "bare ground"
                     Species Code =="W" == Coarse woody debris"
                        Species Code == "F"
                         Species Code == "R"
                         Species Code == T == "thatch
                        TRUE ~ as.character(CommonName)
                        
                      SciName = case_when (
                        Type = "Gound" ~ Common names,
                        TRUE ~ as.character(SciName))
```

```{r}

```


Finish this later: see picture on phone
```{r}
tidying:
  Spring_data_raw_better<- Spring_data_raw >%>
  row_to_names(row_number1)) >%> 
  rename(Transect = 'Transect #', Hit = '# Hits') >%>
  )

```

```{r}

```

```{r}
# the function row_to_names()is from the janitor package
Spring_2020_raw <- Spring_2020_raw %>% row_to_names(row_number(1))

Spring_2020_raw

#same as

row_to_names(Spring_2020_raw, row_number = 1)

```

are those column names okay, or do they need to be changed to be tidy, contain identifying information? use rename() to give your columns better names.
```{r}
Spring_2020_raw <- Spring_2020_raw %>% rename(`Transect#`, `#Hits`)
```

Which rows dont contain observations?
remove them from your dataset using [,] or filter()
#as.Date(42705(colname), origin = ")
```{r}
Spring_2020_raw %>% filter("Date" & Date != "mm/dd/yy")
```
# equals is ==
#Greater than is > or lesser than is <
#equal to or greater is =>, and vice versa =<
#doesnt equal is !=
#and is &
#or is |

What is a factor? regarding anova's, it would be a category.
=======

### 2020 Raw Data Cleaning ###
### 1/12/2023 ###

### Step 1: Install Packages ###

install.packages("tidyverse")
install.packages("readxl")
install.packages("janitor")

library(tidyverse)
library(readxl)
library(janitor)

### Step 2: Import Data ###
## The readxl package allows you to pull data straight from an excel file instead of having to make a CSV file first ##

```{r}
Spring_2020_raw <- read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx")
## multiple sheets are in this file so we need to specify which sheet we want to work with

Spring_2020_raw <- read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1")
## there are a lot of error messages due to the sheet not being cleaned up properly ##

Spring_2020_raw <- read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", skip = 1)
## the skip line allows you to get rid of a column of the dataset. We want to keep the date so we need to get rid of the skip line ##

# Instead lets perhaps skip naming columns and just have everything come in as characters. 

Spring_2020_raw <- read_excel("Spring 2020 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = FALSE)
```

### There is more than one year's worth of data however. We need to import all three and check if they're in the same format.###

```{r}
Spring_2020_headers <- 
  read_excel("Spring 2021 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE)

Spring_2021_headers <- 
  read_excel("Spring 2021 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE)

Spring_2022_headers <- 
  read_excel("Spring 2022 - Percent Cover - Raw Data DNET.xlsx", sheet = "Sheet1", col_names = TRUE)

names(Spring_2020_headers)
names(Spring_2021_headers)
names(Spring_2022_headers)
```

Formatting one data frame - let's check out the checklist!

Which row has the best column names to keep?
Use colnames(data) <- data[rownumber, ] to name your columns after that row.

```{r}

# the function row_to_names() is from the janitor package
Spring_2020_raw <- Spring_2020_raw %>% row_to_names(row_number(1))

Spring_2020_raw
```

Are those column names okay, or do they need to be changed to be tidy, contain identifying information? 
Use rename() to give your columns better names. 
## rename(Name you want it to be = 'current column name') ##
```{r}
Spring_2020_raw <- Spring_2020_raw %>% rename(Transect = `Transect #`, Hits = `# Hits`)
```

### Cleaning Up Data ###
## I am starting with cleaning the block and plot columns ##
## I need to change the values from 1.0 to 1 ##

## Remove informational Row
>>>>>>> ff4e349b5cfda8f66eea084288adc879cc38143f
```{r}

```

<<<<<<< HEAD
=======
# to make it a factor: 
```{r}
glimpse(Spring_2020_raw) # Block is a character, we want it as a factor
Spring_2020_raw<-Spring_2020_raw %>% mutate(Block = as.factor(Block))
Spring_2020_raw<-Spring_2020_raw %>% mutate(Plot = as.factor (Plot))
Spring_2020_raw<-Spring_2020_raw %>% mutate(Treatment = as.factor(Treatment))
Spring_2020_raw<-Spring_2020_raw %>% mutate(Region = as.factor(Region))
Spring_2020_raw<-Spring_2020_raw %>% mutate(Transect = as.factor(Transect))
```

# to make it numeric
```{r}
Spring_2020_raw<-Spring_2020_raw %>% mutate(Distance = as.numeric(Distance))
```

```{r}
view(Spring_2020_raw)
```

# How to turn columns into whole numbers

```{r}
Spring_2020_raw <- Spring_2020_raw %>%
  mutate(Hits=as.numeric(Hits))

glimpse(Spring_2020_raw)
```

## Changing the date ##
```{r}
Spring_2020_raw <- Spring_2020_raw %>% 
  mutate(Date=as.numeric(Date)) %>% mutate(Date = as.Date(Date,origin = "1899-12-30"))
```

### How to check for Typos ###
```{r}
Species.list <- read_csv("DNCodes.csv")

view(Species.list)

### Amy's Code ###
## rename (original column name = 'Name you want')
SpeciesKey <- Full_Species_List_Glossary_reference_only_ %>%
select(1:7) %>%
rename(SpeciesCode = 'Species Code (A-Z)',
        NativeStatus = 'Native/Non-Native',
        CommonName = 'Common Name',
        SciName = 'Full Scientific Name (Genus species)')
```

```{r}
### Amy's Code ###
### Get rid of non data rows
filter(Date ! = "Date" & Date ! = "mm/dd/yy")

### Looking for NA

Spring_data_raw_better$Date %>% unique()

Spring_data_raw_better %>% filter(is.na(Date) == TRUE)

Spring_data_raw_better %>% filter(is.na(Date) == FALSE)

```

```{r}
### Amy's Code ###
## Left Join, Right Join, Full Join and Combining Data Frames##

intermediate <- Spring_data_raw_better %>%
  filter(is.ca(Date) == FALSE) %>%
  full_join(DNCodes, by = "Code") %>%
  left_join(SpeciesKey, by = "SpeciesCode")

```

```{r}
### Amy's Code ###
## When combining multiple data sets how to make sure to inputs the correct values under different conditions 
## If life form is na put in LifeForm. If there is something there then keep the old life form as a character
## case_when (shows up as na(column of interest) == FALSE ~ substitue in that value to the more complete column , TRUE ~ if there is a value then keep the following entry as a character
mutate(
  life.form = case_when(
    is.na(LifeForm) == FALSE ~ LifeForm,
    TRUE ~ as.character(Lifeform)))
  
SciName = case_when(
  Type == "Ground" ~ CommonName,
  TRUE ~ as.character(SciName)
)  


```

## How to delete column
# select(-Column)

## fill(Column) will copy from row directly above 

>>>>>>> ff4e349b5cfda8f66eea084288adc879cc38143f
